---
title: "Project 1"
author: "Abbas"
date: "10/18/2020"
output: html_document


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Load the data 
First load the data activity and check the first five rows to 
have an idea about the data
```{r}
activity<-read.csv("activity.csv")
head(activity)
```

## Cacualte the total steps
To cacualte the total step for each day. First need to group by the date as we want for each day then sum all stpes foe each day. set na.rm = TRUE so that the sum function ignors the missing values

```{r}
data<-activity %>% group_by(date) %>% summarise(StepsSum = sum(steps, na.rm = TRUE))
```

Histgram of step per day
```{r}
b <- aggregate(steps ~ interval, activity, mean)
plot(b$interval,b$steps,type="l",main="Average number of steps by Interval", xlab="Time Intervals", ylab="Average number of steps")
```

The interval that has max steps are computed as:
```{r}
c<-which.max(b$steps)
b[c,]$interval
```

```{r}
hist(data$StepsSum,main="Histogram steps per day", xlab = "Steps per day")
```
Average number of steps per interval 


Mean and median of total number of steps
```{r}
mean(data$StepsSum)

median(data$StepsSum)
```

To find number of total NAs in the activity dataset
```{r}
sum(is.na(activity))
```

Now we need to loop throgut all rows in the dataset and when ever we see a NA we will replace it with the average interval
```{r}
for (i in 1:nrow(activity)) {
     if(is.na(activity$steps[i])) {
         val <- b$steps[which(b$interval == activity$interval[i])]
         activity$steps[i] <- val 
     }
 }
```

Plot histgram of steps per day for the new dataset 
```{r}
data<-activity %>% group_by(date) %>% summarise(StepsSum = sum(steps, na.rm = TRUE))
hist(data$StepsSum)
```

Mean and median of total number of steps
```{r}
mean(data$StepsSum)

median(data$StepsSum)
```

Create a function to check is the date of each row is a weekday or weekend

```{r}
week_day <- function(date_val) {
    wd <- weekdays(as.Date(date_val, '%Y-%m-%d'))
    if  (!(wd == 'Saturday' || wd == 'Sunday')) {
        x <- 'Weekday'
    } else {
        x <- 'Weekend'
    }
    x
}
```

Plot the steps per intervural for weekdays and weekends
```{r}
steps <- aggregate(steps ~ interval+dateDay, activity, mean)
library(ggplot2)
 plt <- ggplot(steps, aes(interval, steps)) +
     geom_line(stat = "identity", aes(colour = dateDay)) +
     theme_gray() +
     facet_grid(dateDay ~ ., scales="fixed", space="fixed") +
     labs(x="Interval", y=expression("No of Steps")) +
     ggtitle("No of steps Per Interval by day type")
print(plt)
```

